---
title: "SIGIZI monitoring"
description: "Aggregate ANC, INC, and PNC monitoring from retained SIGIZI records"
---

```{r}
source("R/render_context.R")
source("R/privacy.R")
source("R/monitoring_data.R")

context <- monitoring_context()
source_summary <- read_public_monitoring_view(
  context,
  "v_sigizi_source_monitoring"
)
stage_summary <- read_public_monitoring_view(
  context,
  "v_sigizi_stage_monitoring"
)
deletion_summary <- read_public_monitoring_view(
  context,
  "v_sigizi_deletion_monitoring"
)
```

::: {.status-banner}
**Aggregate-only section:** the source roles and exclusion rules below follow
the upstream `birth-outcome` pipeline. Live figures appear only when the three
approved monitoring views are deployed and `MONITORING_DATA_SOURCE=bigquery`
is enabled during a controlled render.
:::

## Monitoring pathway

```{mermaid}
flowchart LR
  A[Five cleaned SIGIZI clinical sources] --> B{Pregnancy-specific\ndelection check}
  H[Daftar ibu hamil hapus] --> B
  B -->|Retained records| C[Geography correction]
  B -->|Matched deleted pregnancy| D[Restricted exclusion audit]
  C --> E[Canonical SIGIZI pregnancy episodes]
  E --> F[Aggregate ANC / INC / PNC monitoring views]
  F --> G[Public Quarto site]
```

The deletion registry is an exclusion reference. It is not unioned into ANC,
INC, PNC, pregnancy, or delivery totals.

## Source roles

| SIGIZI source | Pipeline value | Monitoring use | Current boundary |
|---|---|---|---|
| Ibu hamil | `DAFTAR_IBU_HAMIL` | ANC pregnancy/registration evidence | Do not treat a registration row as an ANC encounter unless an ANC date exists |
| Pemeriksaan ANC | `ANC` | ANC examination evidence | Explicit ANC dates support visit monitoring |
| Kohort ibu | `KOHORT_IBU` | INC and delivery evidence | Delivery measures require a valid delivery date |
| Kohort nifas | `KOHORT_NIFAS` | INC evidence and future PNC monitoring | PNC visits/KF stages require a curated visit-level adapter |
| Daftar ibu | `DAFTAR_IBU` | Pregnancy registry and source-presence context | Not an ANC, INC, or PNC event numerator by itself |
| Daftar ibu hamil hapus | `vs_sigizi_bumil_hapus` | Pregnancy-specific exclusion only | Never clinical evidence and never a service numerator |

## ANC, INC, and PNC readiness

::: {.metric-grid}
::: {.metric-card}
### ANC

Uses retained `DAFTAR_IBU_HAMIL` and `ANC` records. Source-record volume,
pregnancies with source evidence, and pregnancies with explicit ANC dates stay
as separate measures.
:::

::: {.metric-card}
### INC

Uses retained `KOHORT_IBU` and `KOHORT_NIFAS` records. Source rows,
pregnancies with delivery evidence, and canonical delivery events must not be
combined into one denominator.
:::

::: {.metric-card}
### PNC

`KOHORT_NIFAS` confirms source availability. KF1-KF4 coverage is intentionally
pending until a curated PNC visit date and visit-stage contract is approved.
:::
:::

```{r}
#| results: asis
render_monitoring_table(
  stage_summary,
  empty_message = paste(
    "Live stage measures are not enabled in this render.",
    "The view contract is ready for BigQuery deployment."
  )
)
```

## Retained source monitoring

Counts below, when enabled, come from `t_sigizi_source_records` after the
deletion gate. Small non-zero counts below five are suppressed in the public
views.

```{r}
#| results: asis
render_monitoring_table(
  source_summary,
  empty_message = "Live retained-source measures are not enabled in this render."
)
```

## Deleted-pregnancy exclusion

A SIGIZI clinical row is excluded only when it matches the same pregnancy in
the deletion registry:

1. Matching usable identifier plus an exactly matching pregnancy anchor; or
2. when an identifier is unavailable on either side, exact normalized name,
   maternal birth date, and pregnancy anchor.

The source anchor is HPHT, falling back to HPL minus 280 days only when HPHT is
missing. Conflicting usable identifiers block the name-based fallback. A later
pregnancy for the same mother remains eligible, and independent EPUS or
hospital evidence is not removed.

```{r}
#| results: asis
render_monitoring_table(
  deletion_summary,
  empty_message = "Live deletion-gate reconciliation is not enabled in this render."
)
```

## Denominator rules

- Raw/retained source records measure source volume.
- `t_sigizi_pregnancy_episode_v3_3` measures canonical SIGIZI pregnancies.
- An ANC registration is not automatically an ANC encounter.
- An INC source record, a pregnancy with delivery evidence, and a canonical
  delivery event are different grains.
- PNC visits and KF1-KF4 cannot be reported until the visit-level adapter is
  explicitly defined.
- All displayed tables are aggregate-only and pass the public-column guard
  before being embedded in the static site.

See the [SIGIZI monitoring contract](https://github.com/oucru-id/sigizi-epus-monitoring/blob/main/docs/sigizi-monitoring-contract.md)
for the SQL objects, deployment order, and validation boundary.
